The globular cluster system of the Auriga simulations

### 2019-08-19 AurigaGCS.tex
Edited Sec. 3 (observations) based on feedback WHT, RG, FM for #1.

\subsection{Milky Way}
\label{sec:milkyway}
\citet[][2010 edition; hereafter H96e10]{1996AJ....112.1487H} provides a
catalogue\footnote{See \url{https://www.physics.mcmaster.ca/Fac_Harris/mwgc.dat}}
of the Milky Way globular cluster system that contains properties of
157 GCs. The authors initially estimated the size of the MW GCS to be 180~$\pm$~10,
thus, their catalogue to be ${\sim}$85\% complete. However, an additional 59 GCs
are claimed to have been discovered by various authors. The total number of GCs
in the MW might add up to 216 with recent estimates now anticipating an additional 
thirty GCs yet to be discovered \citep[e.g.][and references therein]{2018ApJ...863L..38R}.
We still use data from the Harris catalogue, but caution that it may (only) be
53-72\% complete. Specifically, the relevant data fields that we use from H96e10
are the metallicity [Fe/H], the Galactic distance components $X$, $Y$, and $Z$ (in
kpc)\footnote{In a Sun-centered coordinate system: $X$ points toward Galactic
center, $Y$ in direction of Galactic rotation, and $Z$ toward the North Galactic
Pole. We calculate the galactocentric radius $R_{\text{gal}}=\sqrt{(X-R_\odot)^2
+ Y^2 + Z^2}$, assuming the solar radius $R_\odot=8$~kpc.}, and absolute
magnitude in the V-band $M_V$. We use the latter to calculate mass-estimates by
assuming $M_{V,\odot}=4.83$ and a mass to light ratio $M/L_V = 1.7$~M/L$_{\odot}$,
the mean for MW clusters \citep{2005ApJS..161..304M}. We supplement the catalogue
with age-estimates from isochrone fits to stars near the main-sequence turnoff 
in 55 GCs \citep[][hereafter V13]{2013ApJ...775..134V}.

@@

\subsection{Andromeda}
\label{sec:andromeda}
The fifth revision of the revised bologna catalogue (RBC~5, last updated
August, 2012) is the latest edition of three decades of systematically
collecting integrated properties of the globular cluster system of the
Andromeda galaxy \citep[][and references therein]{2004A&A...416..917G}. One
contribution to RBC~5 is the work by \citet[][hereafter C11]{2011AJ....141...61C},
subsequently updated by \citet[][hereafter CR16]{2016ApJ...824...42C}.

C11 and CR16 present a uniform set of spectroscopic observations calibrated
on the Milky Way GCS of the inner $1.6^\circ~({\sim}21)$~kpc that
is believed to be 94\% complete. GCs in the outer stellar halo, up to
$R_{\text{proj}}\sim150$~kpc, are observed in the Pan-Andromeda Archaeological
Survey \citep[PAndAS, ][hereafter H14]{2014MNRAS.442.2165H}, but see also
\citet{2014MNRAS.442.2929V} and \citet{2019MNRAS.484.1756M}. H14 presents the
discovery of 59 new GCs and publishes updates to RBC~5. The work of H14 is
incorporated in the latest public release\footnote{Last revised 23 Sep 2015, see
\url{https://www.cfa.harvard.edu/oir/eg/m31clusters/M31_Hectospec.html}}
of the C11 dataset, further revised by CR16. It seems that CR16 is the most
recent aggregated dataset of M31's GCS that contains properties of interest
for our study as it contains GCs in the inner region and in the outer halo. The
relevant fields in the CR16 dataset that we use are the age, metallicity, and the
mass-estimate\footnote{The authors assumed $M/L_V = 2$ independent of [Fe/H]}.
Radii are calculated from RA and DEC, further discussed in Sec.~\ref{sec:Rgc}.

\subsection{Age-estimates}
The top panel of Figure~\ref{fig:MW-M31-age-Rgc} shows a histogram of the age-estimates
of the $55$ MW GCs in V13 and $88$ GCs in M31 for which age-estimates are available
in CR16. The mean age of the MW GCs in this data set is $11.9 \pm 0.1$~Gyr and the 
dispersion is $0.8$~Gyr. Furthermore, only one of the 55 GC age-estimates is
below $10$~Gyr. The M31 GCS has a mean age of $11.0 \pm 0.2$~Gyr with a dispersion 
of $2.2$~Gyr, and $24$ GCs have age-estimates below $10$~Gyr with a minimum age 
of $4.8$~Gyr. Based on these data, we find that the age distributions of the MW 
and M31 GCS's are not statistically consistent. Moreover, the Milky Way appears 
to host a globular cluster system that is somewhat older than the Andromeda galaxy,
which is somewhat surprising given that M31 is generally considered to be earlier
type than the MW meaning that an older stellar population would naively be expected.
However, we do caution that both data sets are incomplete and the age 
measurements have large uncertainties (of $1-2$~Gyr). On the other hand, the 
magnitude of the uncertainty is insufficient to explain the low-age tail in M31.

\subsection{Radial distribution}
\label{sec:Rgc}
The bottom panel of Figure~\ref{fig:MW-M31-age-Rgc} shows the radial distribution
of the MW and M31 GCS. Both quantities are readily available in H96e10 (assuming
$R_{\odot}=8.0$~kpc), but the galactocentric radius of GCs in M31 is not available
in CR16. Therefore we follow \citet[][Sec.~4.1]{2019arXiv190111229W} to calculate
the projected radius $R_{\text{proj}}$ from the observed positions, adopting M31's
central position from the NASA Extragalactic 
Database\footnote{\url{https://ned.ipac.caltech.edu/}} $(\alpha_0, \, \delta_0) =
(0^{\text{h}}42^{\text{m}}44.35^{\text{s}}, \, +41^{\circ}16'08.63")$
and distance $D_{\text{M31}} = 780$~kpc \citep{2005MNRAS.356..979M,2012ApJ...758...11C}.
We calculate $R_{\text{gal}}$ as `average deprojected distance`
$R_{\text{gal}} = R_{\text{proj}} \times (4/\pi)$.

The solid lines show the distributions using all available data (because the
sky coordinates are known for each GC), while the subset for which age-estimates
are available is indicated using dotted lines. The latter shows a narrower range
of radii than the full data set: it appears that age-estimates are neither
available for the innermost ($<1$~kpc) GCs, nor for those beyond roughly twenty
kpc (the halo GCs). We compare both distributions of the full data set (solid 
lines) and find that the Milky Way has more GCs in the range $1-4$~kpc than M31
(when accounting for the larger number of total GCs in M31). Interestingly, both
distributions show a similar trend for $R_{\text{gal}} > 4$~kpc and host a 
subpopulation of halo GCs. However, both radial distributions are not 
statistically consistent due to substantial differences at intermediate radii.

@@

\subsection{Total GC mass in metallicity-radial space} \label{sec:observations_FeHRgc}
We show the two-dimensional mass-weighted metallicity-radial distribution of the
MW (M31) GCS in the top (bottom) panel of Figure~\ref{fig:observations_FeHRgc}. 
The observations indicate that
no GCs with high metallicities are to be expected at large radii (the three bins
in the upper right corner, both for MW and M31), and relatively few GCs at large
radii in general ($R_{\text{gal}} > 30$~kpc; right column: $11$ GCs or 7.3\% in
the MW and $17$ or 4.6\% in M31). Moreover, the M31 hosts more metal-rich 
([Fe/H]$ > -1$) GCs in each radial bin in comparison to the MW GCSs than what
would be expected when accounting for the fact that Andromeda hosts a larger GC 
system than the MW. Finally, given that both marginalized (i.e. the metallicity 
and radial) distributions are not statistically consistent, we find that the 
two-dimensional distributions are also not consistent. More generally, the GC
systems of the Milky Way and that of Andromeda both differ significantly. We 
compare these observations to the Auriga simulations later on in 
Sec.~\ref{sec:results_FeHRgc}.

@@

\begin{figure}
    \includegraphics[width=\columnwidth]{{FeH_obs_Au4-4_Au4-10_Au4-21}.png}
    \caption{
        \emph{Top}: Metallicity distribution of the MW and M31 GCS. The solid 
        purple (magenta) line shows the GC system of the MW (M31), and we 
        indicate the mean value with an arrow at the bottom of the x-axis. We
        overplot a double Gaussian for the MW GCS using purple dotted lines, taking
        the mean and dispersion from \citep{Harris2001}.
        \emph{Other panels}: three selected Auriga haloes as indicated in the 
        upper right corner. In each of these three panels we show the GG 
        candicates in green, and use a green dotted line for all star particles. 
        We split the GC candidates up into two subpopulations, those that have 
        formed {\it in~situ} (blue), and those that have been accreted (red).
        \label{fig:Au4-10and21_FeH}
    }
\end{figure}
We select \mbox{Au4-10}\footnote{The nomenclature is `Au' for Auriga, followed by the
resolution level (4) and halo number (10, indicating which set of initial
conditions was used the run).}, \mbox{Au4-21}, and \mbox{Au4-4} to show the
distributions of three individual simulation runs that we consider representative
for specific behaviour, and to highlight that different runs could give rise to 
different distributions. We plot the age-selected GC candidates in green, those
that are accreted in red, and the {\it in~situ} subpopulation in blue. The top 
half of the left figure shows the MW (M31) GC system in purple (magenta), where 
we overplot a double Gaussian for the MW GCS (the purple dashed lines) by 
adopting literature values of the mean $\mu$ and standard deviation $\sigma$ of 
the metal-rich and metal-poor subpopulations \citep[][p. 38]{Harris2001}.

>>>>

\subsection{Milky Way}
\label{sec:milkyway}
\citet[][2010 edition; hereafter H96e10]{1996AJ....112.1487H} provides a
catalogue\footnote{See \url{https://www.physics.mcmaster.ca/Fac_Harris/mwgc.dat}}
of the Milky Way globular cluster system that contains properties of 157 GCs. 
The authors initially estimated the number of GCs in the MW GCS to be 180~$\pm$~10,
thus, their catalogue to be ${\sim}$85\% complete. However, an additional 59 GCs
are claimed to have been discovered by various authors. The total number of GCs
in the MW might add up to 216 with recent estimates now anticipating an additional 
thirty GCs yet to be discovered \citep[e.g.][and references therein]{2018ApJ...863L..38R}.
We still use data from the Harris catalogue, but caution that it may (only) be
53-72\% complete. Specifically, the relevant data fields that we use from H96e10
are the metallicity [Fe/H], the Galactic distance components $X$, $Y$, and $Z$ (in
kpc)\footnote{In a Sun-centered coordinate system: $X$ points toward Galactic
center, $Y$ in direction of Galactic rotation, and $Z$ toward the North Galactic
Pole. We calculate the galactocentric radius $R_{\text{gal}}=\sqrt{(X-R_\odot)^2
+ Y^2 + Z^2}$, assuming the solar radius $R_\odot=8$~kpc.}, and absolute
magnitude in the V~band $M_V$. We use the latter to calculate mass-estimates by
assuming $M_{V,\odot}=4.83$ and a mass to light ratio $M/L_V = 1.7$~M/L$_{\odot}$,
the mean for MW clusters \citep{2005ApJS..161..304M}. We supplement the catalogue
with age-estimates from isochrone fits to stars near the main-sequence turnoff 
in 55 GCs \citep[][hereafter V13]{2013ApJ...775..134V}.

@@

\subsection{Andromeda}
\label{sec:andromeda}
The fifth revision of the revised Bologna catalogue (RBC~5, last updated
August, 2012) is the latest edition of three decades of systematically
collecting integrated properties of the globular cluster system of the
Andromeda galaxy \citep[][and references therein]{2004A&A...416..917G}. One
contribution to RBC~5 is the work by \citet[][hereafter C11]{2011AJ....141...61C},
subsequently updated by \citet[][hereafter CR16]{2016ApJ...824...42C}.

C11 and CR16 present a uniform set of spectroscopic observations calibrated
on the Milky Way GCS of the inner $1.6^\circ~({\sim}21)$~kpc that
is believed to be 94\% complete. GCs in the outer stellar halo, up to
$R_{\text{proj}}\sim150$~kpc, are observed in the Pan-Andromeda Archaeological
Survey \citep[PAndAS, ][hereafter H14]{2014MNRAS.442.2165H}, but see also
\citet{2014MNRAS.442.2929V} and \citet{2019MNRAS.484.1756M}. H14 presents the
discovery of 59 new GCs and publishes updates to RBC~5. The work of H14 is
incorporated in the latest public release\footnote{Last revised 23 Sep 2015, see
\url{https://www.cfa.harvard.edu/oir/eg/m31clusters/M31_Hectospec.html}}
of the C11 dataset, further revised by CR16. We use CR16 because it is the most
recent aggregated dataset of M31's GCS that contains properties of interest
for our study and it includes GCs in the inner region and in the outer halo. The
relevant fields in the CR16 dataset that we use are the age, metallicity, and the
mass-estimate\footnote{The authors assumed $M/L_V = 2$ independent of [Fe/H]}.
Radii are calculated from RA and DEC, further discussed in Sec.~\ref{sec:Rgc}.

\subsection{Age-estimates}
The top panel of Figure~\ref{fig:MW-M31-age-Rgc} shows a histogram of the age-estimates
of the $55$ MW GCs in V13 and $88$ GCs in M31 for which age-estimates are available
in CR16. The mean age of the MW GCs in this data set is $11.9 \pm 0.1$~Gyr and the 
dispersion is $0.8$~Gyr. Furthermore, only one of the 55 GC age-estimates is
below $10$~Gyr. The M31 GCS has a mean age of $11.0 \pm 0.2$~Gyr with a dispersion 
of $2.2$~Gyr, and $24$ GCs have age-estimates below $10$~Gyr with a minimum age 
of $4.8$~Gyr. Based on these data, we find that the age distributions of the MW 
and M31 GCS's are not statistically consistent. Moreover, the Milky Way appears 
to host a globular cluster system that is somewhat older than the Andromeda galaxy,
which is somewhat surprising given that M31 is generally considered to be earlier
type than the MW meaning that an older stellar population would be naively expected.
However, we do caution that both data sets are incomplete and the age 
measurements have large uncertainties (of $1-2$~Gyr). On the other hand, the 
magnitude of the uncertainty is insufficient to explain the low-age tail in M31.

\subsection{Radial distribution}
\label{sec:Rgc}
The bottom panel of Figure~\ref{fig:MW-M31-age-Rgc} shows the radial distribution
of the MW and M31 GCS. Both quantities are readily available in H96e10 (assuming
$R_{\odot}=8.0$~kpc), but the galactocentric radius of GCs in M31 is not available
in CR16. Therefore, we follow \citet[][Sec.~4.1]{2019arXiv190111229W} to calculate
the projected radius $R_{\text{proj}}$ from the observed positions, adopting M31's
central position from the NASA Extragalactic 
Database\footnote{\url{https://ned.ipac.caltech.edu/}} $(\alpha_0, \, \delta_0) =
(0^{\text{h}}42^{\text{m}}44.35^{\text{s}}, \, +41^{\circ}16'08.63")$
and distance $D_{\text{M31}} = 780$~kpc \citep{2005MNRAS.356..979M,2012ApJ...758...11C}.
We calculate $R_{\text{gal}}$ as `average deprojected distance`
$R_{\text{gal}} = R_{\text{proj}} \times (4/\pi)$.

The solid lines show the distributions using all available data (because the
sky coordinates are known for each GC), while the subset for which age-estimates
are available is indicated using dotted lines. The latter shows a narrower range
of radii than the full data set: it appears that few age-estimates are available 
for the innermost ($<1$~kpc) GCs, and none for those beyond roughly twenty
kpc (the halo GCs). We compare both distributions of the full data set (solid 
lines) and find that the Milky Way has more GCs in the range $1-4$~kpc than M31
(when accounting for the larger number of total GCs in M31). Interestingly, both
distributions show a similar trend for $R_{\text{gal}} > 4$~kpc and host a 
subpopulation of halo GCs. However, the two radial distributions are not statistically
consistent with being drawn from the same underlying distribution due to substantial 
differences at intermediate radii.

@@

\subsection{Total GC mass in metallicity-radial space} \label{sec:observations_FeHRgc}
We show the two-dimensional mass-weighted metallicity-radial distribution of the
MW (M31) GCS in the top (bottom) panel of Figure~\ref{fig:observations_FeHRgc}. 
The observations indicate that
no GCs with high metallicities are to be expected at large radii (the three bins
in the upper right corner, both for MW and M31), and relatively few GCs at large
radii in general ($R_{\text{gal}} > 30$~kpc; right column: $11$ GCs or 7.3\% in
the MW and $17$ or 4.6\% in M31). Moreover, the M31 hosts more metal-rich 
([Fe/H]$ > -1$) GCs in each radial bin in comparison to the MW GCSs than what
would be expected when accounting for the fact that Andromeda hosts a larger GC 
system than the MW. Finally, given that both marginalized (i.e. the metallicity 
and radial) distributions are not statistically consistent, we find that the 
two-dimensional distributions are also not consistent. More generally, the GC
system of the MW differs significantly from that of M31. We compare these 
observations to the Auriga simulations later on in Sec.~\ref{sec:results_FeHRgc}.

@@

\begin{figure}
    \includegraphics[width=\columnwidth]{{FeH_obs_Au4-4_Au4-10_Au4-21}.png}
    \caption{
        \emph{Top}: Metallicity distribution of the MW and M31 GCS. The solid 
        purple (magenta) line shows the GC system of the MW (M31), and we 
        indicate the mean value with an arrow at the bottom of the x-axis. We
        overplot a double Gaussian for the MW GCS using purple dotted lines, taking
        the mean and dispersion from \citep{Harris2001}.
        \emph{Other panels}: three selected Auriga haloes as indicated in the 
        upper right corner. In each of these three panels we show the GG 
        candicates in green, and use a green dotted line for all star particles. 
        We split the GC candidates up into two subpopulations, those that have 
        formed {\it in~situ} (blue), and those that have been accreted (red).
        \label{fig:Au4-10and21_FeH}
    }
\end{figure}
We select \mbox{Au4-10}\footnote{The nomenclature is `Au' for Auriga, followed by the
resolution level (4) and halo number (10, indicating which set of initial
conditions was used).}, \mbox{Au4-21}, and \mbox{Au4-4} to show the
distributions of three individual simulation runs that we consider representative
for specific behaviour, and to highlight that different runs could give rise to 
different distributions. We plot the age-selected GC candidates in green, those
that are accreted in red, and the {\it in~situ} subpopulation in blue. The top 
half of the left figure shows the MW (M31) GC system in purple (magenta), where 
we overplot a double Gaussian for the MW GCS (the purple dashed lines) by 
adopting literature values of the mean $\mu$ and standard deviation $\sigma$ of 
the metal-rich and metal-poor subpopulations \citep[][p. 38]{Harris2001}.
